Safe Memory Regions for Big Data Processing

### 2016-06-01 paper/type-inference.tex
Type inference

@@

Region type templates are FGJ types extended with fresh region
variables ($\rho$) and predicate variables ($\varphi$) to denote
unknown region annotations and region constraints, respectively. For
instance, if a variable $x$ has type $\ObjZ$ in FGJ, its region type
template is of form $\ObjZ\inang{\rho_0}$, where $\rho_0$ is a fresh
region variable. Likewise, given the region-annotated definition of
\C{Pair} class from \S\ref{sec:fb-syntax} a region type template for a
method with FGJ type $\C{Pair}\inang{\C{A},\C{B}} \rightarrow \C{A}$
is \footnote{In our exposition, we assume that classes $\C{A}$ and
$\C{B}$ are trivial subclasses of $\ObjZ$ with no fields/methods. Like
$\ObjZ$, they accept one region parameter - the allocation region of
their objects.}\footnote{We abuse arrow notation to also represent
types of methods, but unlike function types, there is no allocation
region annotation atop the arrow in a method type.}
$\inang{\rhoalloc_0,\rho_1,\rho_2,\rho_3 \,|\, \varphi_0}
\C{Pair}\inang{\C{A},\C{B}} \inang{\rho_1,\rho_2,\rho_3} \rightarrow
\unitZ$, where $\rhoalloc_0$ and $\rho_{1-3}$ are fresh region
variables, and $\varphi_0$ is a fresh predicate variable denoting
unknown constraints over $\rhoalloc_0$ and $\rho_{1-3}$.  Region type
template of a type variable $\C{a}$ is $\C{a}@\rho$, where $\rho$ is
fresh. For a class, a template is computed in two steps. In the first
step, we templatize the types of all its fields, constructor
arguments, and arguments and return values of all its methods, along
with the type of its superclass. In the second step, we generalize all
the free region variables occuring in the templates computed in the
first step as region parameters of the class. Finally, we add a fresh
allocation region parameter ($\rhoalloc$) to the list of parameters,
and introduce a new predicate variable ($\varphi$) to denote unknown
constraints on region parameters. For example, consider the standard
FGJ definition of $\C{Pair}\inang{\C{a},\C{b}}$ class, where $\C{a}$
and $\C{b}$ are the types of $\C{fst}$ and $\C{snd}$ fields,
respectively. It can be templatized as following:
\begin{center}
\begin{codejava}
class Pair$\inang{a\extends\ObjZ, b\extends\ObjZ}$
          $\inang{\rhoalloc_0,\rho_{0-4}\,|\,\varphi_0} \extends \ObjZ\inang{\rho_4}$ {
  $a@\rho_0$ fst;
  $b@\rho_1$ snd;
  Pair($a@\rho_2$ fst, $b@\rho_3$ snd) { $...$ }
  $...$
}
\end{codejava}
\end{center}
% \footnote{In reality, the generalized region variables only come
% from the type templates of class's fields and superclass. Method
% types do not contain any free region variables.}
The template elides $\C{Pair}$'s methods, whose region type templates
contain no free variables. Among the region parameters of the class
template, $\rho_{0-4}$ are obtained by generalizing free region
variables in the types of its class's fields, constructor arguments,
and its superclass type.  The remaining parameter ($\rhoalloc_0$) is a
fresh region variable denoting the allocation region argument. Fresh
predicate variable $\varphi_0$ denotes unknown constraints over
$\rhoalloc_0$ and $\rho_{0-4}$ that need to hold for template to be a
well-formed region-annotated class definition in $\FB$.

@@

Our constraint generation algorithm generates three kinds of
constraints:
\begin{itemize}
\item Well-formedness constraints of form $\rho \in \rhoenv$,
restricting the domain of unification for a region variable ($\rho$)
to the set ($\rhoenv$) of regions in scope,
\item Well-formedness constraints of form
$\tywf{\rhoenv}{\varphi}$, restricting the domain of a predicate
variable ($\varphi$) to the set of all possible constraint formulas
over region variables ($\rhoenv$) in scope, and
\item Validity constraints of form $\isvalid{\phicx}{\phi}$, where
$\phicx$ and $\phi$ are region constraints (Fig.~\ref{fig:fb-syntax})
extended with predicate variables and \emph{pending
substitutions}\footnote{we borrowed this terminology
from~\cite{ltpldi08}}:
\begin{smathpar}
\begin{array}{lcl}
\phicx,\phi & \coloneqq & true \ALT \rho \outlives \rho \ALT \rho = \rho 
    \ALT F(\varphi) \ALT \phi \conj \phi\\
F & \coloneqq & \cdot \ALT [\rho/\rho]F \\
\end{array}
\end{smathpar}
\end{itemize}
A pending substitution ($F$) is a substitution function over region
variables/identifiers. They represent the substitutions that need to
be carried out when a predicate variable ($\varphi$) is replaced by a
concrete formula in a validity constraint. For instance, in the
validity constraint $\isvalid{\rgn_1 \outlives
\rgn_2}{[\rgn_1/\rho_1][\rgn_2/\rho_2]\varphi}$, pending substitution
is $[\rgn_1/\rho_1][\rgn_2/\rho_2]$. Any concrete formula (call it
$\phisol$) over variables $\rho_1$ and $\rho_2$ is a solution to
$\varphi$ if and only the formula obtained by substituting $\rgn_1$
and $\rgn_2$ for $\rho_1$ and $\rho_2$ (resp.) in $\phisol$ is
deducible from $\rgn_1 \outlives \rgn_2$.

% Validity constraints primarily result from checking if a region type
% template of form $B\inang{\tbar}\inang{\rhoalloc,\rhobar}$ is
% well-formed (i.e., $\typeOk$). The corresponding well-formedness rule
% (Fig.~\ref{fig:fb-staticsem}) requires that the instantiation of $B$'s
% formal region parameters with actual region arguments satisfy its
% precondition, where actuals are subsituted for formals. Since $B$'s
% precondition could be unknown, the requirement results in a in a
% validity constraint with predicate variables and pending
% substitutions. Other sources of validity constraints are calls to
% region-polymorphic methods, where similar situation arises. 
In general, validity constraints generated by our algorithm assume one
of the following two forms:
\begin{center}
\( \isvalid{\phictxt \conj \varphi_i}
           {\phicstr} \qquad
   \isvalid{\phictxt \conj \varphi_i}
           {F_j(\varphi_j)}
\)
\end{center}
where $\varphi_i$ and $\varphi_j$ denote the unknown preconditions of some
methods or constructors.
%
%% If the constraint is generated while checking the well-formedness of a
%% type or elaborating an expression, then $\varphi_j$'s denote the unknown
%% preconditions of classes and methods that were used in that type or expression.
%
Each use of a (region-polymorphic) class or a method may instantiate region
parameters differently, resulting in a different pending substitution
($F_j$). Formulas $\phictxt$ and $\phicstr$ are concrete, i.e., free
of predicate variables and pending substitutions. While $\phictxt$
captures relationships that are \emph{known} to hold between concrete
region identifiers (i.e., $\rgn$'s) when the constraint was generated,
$\phicstr$ captures relationships that are \emph{required} to hold
among region varibles (i.e., $\rho$'s), or relationships between
region variables and identifiers.
%
Each region variable occuring in $\phicstr$ has an associated well-formedness
constraint, which specifies its unification domain. The unification domain of a
constraint is the union of unification domains of all region variables
occuring in the constraint.

@@

\paragraph{Constraints Example 2} Let us add to the \C{Pair} class a
contrived method $\C{alt}$ that accepts a \C{Region} object \C{r}, a
\C{Pair<A,A>} object \C{q}, and an \C{A} object \C{y}. It assigns
\C{y} to \C{fst} and \C{snd} fields of \C{q}, and calls itself
recursively with the same region, a new \C{Pair} object allocated in a
local region, and an \C{A} object referred by the \C{snd} field of the
pair inside the region. \C{alt} never terminates.  Elaboration phase
elaborates the method to the following region-annotated
definition\footnote{In reality, elaboration uses new region variables
as parameters to the constructor and method calls, and then generates
constraints that unify them with actuals. In our examples, to avoid
clutter due to trivial constraints,we coalesced both steps and show
the actuals instead.}(The original definition of \C{alt} can be
obtained by erasing all the region annotations from the elaborated
version):
% \begin{codejava}
% unit alt<$\rhoalloc_2$,$\rho_{6-10}$ | $\varphi_2$>(Pair<A,A><$\rho_6$,$\rho_7$,$\rho_8$> p,
%                           A<$\rho_9$> x, A<$\rho_{10}$> y) {
%   p.fst = x; p.snd = y; 
%   this.alt<$\rhoalloc_2$,$\rho_7$,$\rho_8$,$\rho_{10}$,$\rho_9$>(A,y,x);
% }
% \end{codejava}
\begin{codejava}
unit alt<$\rhoalloc_2$,$\rho_{6-9}\,$|$\,\varphi_2$>(Region<Pair<A,A>><$\toprgn$> r, 
                Pair<A,A><$\rho_{6-8}$> q, A<$\rho_{9}$> y) {
  q.fst := y; q.snd := y; 
  open r as $\rgn_0$ withroot p in
    letregion $\rgn_1$ in
      let x = new Pair<A,A><$\rgn_1$,$\rgn_0$,$\rgn_0$>
                      (p.fst,p.fst) in
        alt<$\rgn_{1}$,$\rgn_{1}$,$\rgn_{0}$,$\rgn_{0}$,$\rgn_{0}$>(r,x,p.snd)
}
\end{codejava}
% Note that, to avoid clutter, we have already resolved appropriate
% region arguments to the recursive call, instead of introducing new
% region variables and generating equality constraints on them. 
Constraints generated during the elaboration are shown below
(let $\rhoenv_2 = \{\rhoalloc_0,\rho_{0-4},\rhoalloc_2,\rho_{6-9}\}$ ):
\begin{smathpar}
\begin{array}{l}
\csid{7} \tywf{\rhoenv_2}{\varphi_2}\qquad
\csid{8} \isvalid{\varphi_1 \conj \varphi_2}
    {\rho_7 \outlives \rho_6 \conj \rho_8 \outlives \rho_6} \\
\csid{9} \isvalid{\varphi_1 \conj \varphi_2}{\rho_7 = \rho_9} \qquad
\csid{10} \isvalid{\varphi_1 \conj \varphi_2}{\rho_{8} = \rho_9} \\
\csid{11} \isvalid{\varphi_1 \conj \varphi_2 \conj \rgn_0 \outlives
\rgn_1} 
    {[\rgn_1/\rhoalloc_2][\rgn_1/\rho_6][\rgn_0/\rho_{7-9}]\varphi_2}
\end{array}
\end{smathpar}

@@

\subsection{Constraint Solving}
\label{sec:fb-constraintsolving}

@@

Solving the set ($C$) of constraints entails finding an
assignment for $C$'s subject ($\varphi$), and also all the region
variables ($\rhobar$) that occur free in $C$, such that the solution
satisfies well-formedness constraints on $\varphi$ and $\rhobar$. To
simplify presentation, we think of $C$ as being parameterized on
$\varphi$ and $\rhobar$, and write it as $C\lbrack \varphi,\rhobar
\rbrack$. We now formalize the constraint satisfaction problem, and
its solution.

\begin{definition}
\emph{(\textbf{Constraint Satisfaction Problem (CSP)})} A constraint
satisfaction problem is a tuple $\CSP$, where $C\lbrack
\varphi,\rhobar \rbrack$ is a set of validity constraints, where
$i$'th validity constraint assumes one of the following forms:
\begin{center}
\(
    \isvalid{\phictxt^{i} \conj \varphi}
            {\phicstr^{i}}\qquad
    \isvalid{\phictxt^{i} \conj \varphi}
            {F_i(\varphi)}
\)
\end{center}
$\bar{\rhoenv}$ are the unification domains for $\rhobar$. We call the
union of all unification domains ($\bigcup\bar{\rhoenv}$) as the
unification domain ($\rhoenv$) of the CSP.  The solution to the
constraint satisfaction problem is a pair $(\substFn,\phisol)$, where
$\substFn$ is a map from $\rhobar$ to $\rhoenv$ such that
$\substFn(\rho_j) \in \rhoenv_j$, for every $j$, and $\phisol$ is a
constraint formula such that:
\begin{itemize}
\item $\phisol$ is well-formed under $\rhoenv_{\varphi}$ (i.e.,
$\tywf{\rhoenv_{\varphi}}{\phisol}$).
\item Every sequent in $C\lbrack (\phisol,\substFn(\rhobar)) \rbrack$
is valid.
% \begin{center}
% \(
%   \isvalid{\phictxt^{i} \conj \bigwedge_j(\rho_j = \substFn(\rho_j)) \conj \phisol}
%           {\phicstr^{i} \conj  F_i(\phisol))}
% \)
% \end{center}
\item $\phisol$ is maximally weak. That is, $\nexists \phisol'$ such
that $\phisol'$ is well-formed, $C\lbrack (\phisol',\substFn(\rhobar))
\rbrack$ is valid, and $\phisol'$ is strictly weaker than $\phisol$
(i.e., $\isvalid{\cdot}{\phisol \Rightarrow \phisol'}$ but
$\isnotvalid{\cdot}{\phisol' \Rightarrow \phisol}$).
\end{itemize}
\end{definition}

Our approach is based on the observation that $\FB$ does not admit
null values or uninitialized variables, thus forcing every region
variable to be unified with some concrete region. The constraint
formula $\Phics$ captures all such unification constraints on
$\rhobar$. Since the set of all concrete region identifiers is the
unification domain ($\rhoenv$) of CSP, this means for every $\rho_i$
with a well-formedness constraint as $\rho_i \in \rhoenv_i$, there
exists a $\rgn \in \rhoenv$ such that and $\isvalid{\Phicx \conj
\Phics}{\rho = \rgn}$. However, $\rgn$ may not belong to $\rhoenv_i$,
in which case the well-formedness constraint on $\rho_i$ is not
satisfied, and constraint solving must fail.  Therefore, there exists
a unique assignment $\substFn$ such that $c\lbrack
\phi_{sol},\substFn(\rhobar)$ is satisfied, regardless of
$\phi_{sol}$.

\item Case $\tynwf{\rhoenv_{\varphi}}{\phi}$: In this case, $\phi$
contains at least one equality or outlives constraint on two region
identifiers, $\rgn_i$ and $\rgn_j$, where (a). $\rgn_i,\rgn_j \in
\rhoenv - \rhoenv_{\varphi}$, or (b). $\rgn_i \in \rhoenv -
\rhoenv_{\varphi}$ and $\rgn_j \in \rhoenv_{\varphi}$, such that the
constraint is not implied by $\Phicx$ (if it is implied, then $\phi$
is not maximally weak). Let us denote such constraint on $\rgn_i$ and
$\rgn_j$ as $\phi_{ij}$. Now, let us consider a solution $\phisol$ to
$\varphi$, which means that $\isvalid{\Phicx \conj \phisol}{\Phics'}$.
Since $\isvalid{\Phicx}{\phi \Leftrightarrow \Phics'}$, we have
$\isvalid {\Phicx \conj \phisol}{\phi}$. Since
$\isvalid{\phi}{\phi_{ij}}$, we have $\isvalid {\Phicx \conj
\phisol}{\phi_{ij}}$, although $\isnotvalid{\Phicx}{\phi_{ij}}$. But
this is impossible. To see why, recall that all the constraints on
identifiers in $\rhoenv_{\varphi}$ occuring in $\Phicx$ are subsumed
by $(\rhoenv - \rhoenv_{\varphi}) \outlives \rhoenv_{\varphi}$.
Therefore, it is impossible to derive $\phi_{ij}$ from $\Phicx$ by
only adding constraints on $\rgn \in \rhoenv_{\varphi}$. Hence, such a
solution $\phisol$ cannot exist.
\end{itemize}
The above discussion hints at an algorithm to compute a solution to
$\varphi$: find a maximally weak $\phisol$ such that $\isvalid{\Phicx}
{\phisol \Leftrightarrow \Phics'}$. If $\tywf{\rhoenv_{\varphi}}
{\phisol}$, then $\phisol$ is the solution. Otherwise, there is no
solution to $\varphi$. We now describe a graph-based algorithm to
compute maximally weak $\phisol$.

>>>>

\textbf{TODO: Mention that constraints are in the theory of partial orders?}

@@

Region type templates are FGJ types extended with fresh region
variables ($\rho$) and predicate variables ($\varphi$) to denote
unknown region annotations and region constraints, respectively. For
instance, if a variable $x$ has type $\ObjZ$ in FGJ, its region type
template is of form $\ObjZ\inang{\rho_0}$, where $\rho_0$ is a fresh
region variable. Likewise, given the region-annotated definition of
\C{Pair} class from \S\ref{sec:fb-syntax} a region type template for a
method with FGJ type $\C{Pair}\inang{\C{A},\C{B}} \rightarrow \C{A}$
is \footnote{In our exposition, we assume that classes $\C{A}$ and
$\C{B}$ are trivial subclasses of $\ObjZ$ with no fields/methods. Like
$\ObjZ$, they accept one region parameter - the allocation region of
their objects.}\footnote{We abuse arrow notation to also represent
types of methods, but unlike function types, there is no allocation
region annotation atop the arrow in a method type.}
$\inang{\rhoalloc_0,\rho_1,\rho_2,\rho_3 \,|\, \varphi_0}
\C{Pair}\inang{\C{A},\C{B}} \inang{\rho_1,\rho_2,\rho_3} \rightarrow
\unitZ$, where $\rhoalloc_0$ and $\rho_{1-3}$ are fresh region
variables, and $\varphi_0$ is a fresh predicate variable denoting
unknown constraints over $\rhoalloc_0$ and $\rho_{1-3}$.  Region type
template of a type variable $\C{a}$ is $\C{a}@\rho$, where $\rho$ is
fresh.
\textbf{TODO: What about bounded types?}
For a class, a template is computed in two steps. In the first
step, we templatize the types of all its fields, constructor
arguments, and arguments and return values of all its methods, along
with the type of its superclass. In the second step, we generalize all
the free region variables occuring in the templates computed in the
first step as region parameters of the class. Finally, we add a fresh
allocation region parameter ($\rhoalloc$) to the list of parameters,
and introduce a new predicate variable ($\varphi$) to denote unknown
constraints on region parameters. For example, consider the standard
FGJ definition of $\C{Pair}\inang{\C{a},\C{b}}$ class, where $\C{a}$
and $\C{b}$ are the types of $\C{fst}$ and $\C{snd}$ fields,
respectively. It can be templatized as following:
\begin{center}
\begin{codejava}
class Pair$\inang{a\extends\ObjZ, b\extends\ObjZ}$
          $\inang{\rhoalloc_0,\rho_{0-4}\,|\,\varphi_0} \extends \ObjZ\inang{\rho_4}$ {
  $a@\rho_0$ fst;
  $b@\rho_1$ snd;
  Pair($a@\rho_2$ fst, $b@\rho_3$ snd) { $...$ }
  $...$
}
\end{codejava}
\end{center}
% \footnote{In reality, the generalized region variables only come
% from the type templates of class's fields and superclass. Method
% types do not contain any free region variables.}
The template elides $\C{Pair}$'s methods, whose region type templates
contain no free variables. Among the region parameters of the class
template, $\rho_{0-4}$ are obtained by generalizing free region
variables in the types of its class's fields, constructor arguments,
and its superclass type.  The remaining parameter ($\rhoalloc_0$) is a
fresh region variable denoting the allocation region argument. Fresh
predicate variable $\varphi_0$ denotes unknown constraints over
$\rhoalloc_0$ and $\rho_{0-4}$ that need to hold for template to be a
well-formed region-annotated class definition in $\FB$.

\textbf{TODO: discuss base class parameters}

@@

Our constraint generation algorithm generates the following kinds of constraints:
\begin{itemize}
\item Well-formedness constraints of form $\rho \in \rhoenv$,
restricting the domain of unification for a region variable ($\rho$)
to the set ($\rhoenv$) of regions in scope,
\item Well-formedness constraints of form
$\tywf{\rhoenv}{\varphi}$, restricting the domain of a predicate
variable ($\varphi$) to the set of all possible constraint formulas
over region variables ($\rhoenv$) in scope, and
\item Validity constraints of form $\isvalid{\phictxt \conj \varphi_i} {\phicstr}$
where $\varphi_i$ is a predicate variable (representing the precondition of a
method to be determined), $\phicstr$ is a region constraint that is \emph{required}
to hold at a particular program point (within the method), and $\phictxt$ is
a region constraint that is \emph{known} to hold at that program point.
Each $\phi$ is a region constraint over region variables and region parameters:
\begin{smathpar}
\begin{array}{lcl}
\phi & \coloneqq & true \ALT \rho \outlives \rho \ALT \rho = \rho \ALT \phi \conj \phi \\
\end{array}
\end{smathpar}
%
%% Formulas $\phictxt$ and $\phicstr$ are concrete, i.e., free
%% of predicate variables and pending substitutions. While $\phictxt$
%% captures relationships that are \emph{known} to hold between concrete
%% region identifiers (i.e., $\rgn$'s) when the constraint was generated,
%% $\phicstr$ captures relationships that are \emph{required} to hold
%% among region varibles (i.e., $\rho$'s), or relationships between
%% region variables and identifiers.
%
\item Validity constraints of the form $\isvalid{\phictxt \conj \varphi_i} {F_j(\varphi_j)}$
generated by an invocation of a method with precondition $\varphi_j$ (where $\phictxt$ and
$\varphi_i$ are as above).
Here, $F$ denotes \emph{pending substitutions}\footnote{we borrowed this terminology from~\cite{ltpldi08}}:
it serves to bind formal region parameters in $\varphi_j$ to the actual region parameters
at a particular invocation site.
\begin{smathpar}
\begin{array}{lcl}
F & \coloneqq & \cdot \ALT [\rho/\rho]F \\
\end{array}
\end{smathpar}
\end{itemize}

A pending substitution ($F$) is a substitution function over region
variables/identifiers. They represent the substitutions that need to
be carried out when a predicate variable ($\varphi$) is replaced by a
concrete formula in a validity constraint. For instance, in the
validity constraint $\isvalid{\rgn_1 \outlives
\rgn_2}{[\rgn_1/\rho_1][\rgn_2/\rho_2]\varphi}$, pending substitution
is $[\rgn_1/\rho_1][\rgn_2/\rho_2]$. Any concrete formula (call it
$\phisol$) over variables $\rho_1$ and $\rho_2$ is a solution to
$\varphi$ if and only the formula obtained by substituting $\rgn_1$
and $\rgn_2$ for $\rho_1$ and $\rho_2$ (resp.) in $\phisol$ is
deducible from $\rgn_1 \outlives \rgn_2$.

% Validity constraints primarily result from checking if a region type
% template of form $B\inang{\tbar}\inang{\rhoalloc,\rhobar}$ is
% well-formed (i.e., $\typeOk$). The corresponding well-formedness rule
% (Fig.~\ref{fig:fb-staticsem}) requires that the instantiation of $B$'s
% formal region parameters with actual region arguments satisfy its
% precondition, where actuals are subsituted for formals. Since $B$'s
% precondition could be unknown, the requirement results in a in a
% validity constraint with predicate variables and pending
% substitutions. Other sources of validity constraints are calls to
% region-polymorphic methods, where similar situation arises.

%
%% If the constraint is generated while checking the well-formedness of a
%% type or elaborating an expression, then $\varphi_j$'s denote the unknown
%% preconditions of classes and methods that were used in that type or expression.
%
% Each use of a (region-polymorphic) class or a method may instantiate region
% parameters differently, resulting in a different pending substitution
% ($F_j$).

Each region variable occuring in $\phicstr$ has an associated well-formedness
constraint, which specifies its unification domain. The unification domain of a
constraint is the union of unification domains of all region variables
occuring in the constraint.

@@

\paragraph{Constraints Example 2} Let us add to the \C{Pair} class a
contrived method $\C{alt}$ that accepts a \C{Region} object \C{r}, a
\C{Pair<A,A>} object \C{q}, and an \C{A} object \C{y}. It assigns
\C{y} to \C{fst} and \C{snd} fields of \C{q}, and calls itself
recursively with the same region, a new \C{Pair} object allocated in a
local region, and an \C{A} object referred by the \C{snd} field of the
pair inside the region. \C{alt} never terminates.  Elaboration phase
elaborates the method to the following region-annotated
definition\footnote{In reality, elaboration uses new region variables
as parameters to the constructor and method calls, and then generates
constraints that unify them with actuals. In our examples, to avoid
clutter due to trivial constraints,we coalesced both steps and show
the actuals instead.}(The original definition of \C{alt} can be
obtained by erasing all the region annotations from the elaborated
version):
% \begin{codejava}
% unit alt<$\rhoalloc_2$,$\rho_{6-10}$ | $\varphi_2$>(Pair<A,A><$\rho_6$,$\rho_7$,$\rho_8$> p,
%                           A<$\rho_9$> x, A<$\rho_{10}$> y) {
%   p.fst = x; p.snd = y; 
%   this.alt<$\rhoalloc_2$,$\rho_7$,$\rho_8$,$\rho_{10}$,$\rho_9$>(A,y,x);
% }
% \end{codejava}
\begin{codejava}
unit alt<$\rhoalloc_2$,$\rho_{6-9}\,$|$\,\varphi_2$>(Region<Pair<A,A>><$\toprgn$> r, 
                Pair<A,A><$\rho_{6-8}$> q, A<$\rho_{9}$> y) {
  q.fst := y; q.snd := y; 
  open r as p@$\rgn_0$ in
    letregion $\rgn_1$ in
      let x = new Pair<A,A><$\rgn_1$,$\rgn_0$,$\rgn_0$>
                      (p.fst,p.fst) in
        alt<$\rgn_{1}$,$\rgn_{1}$,$\rgn_{0}$,$\rgn_{0}$,$\rgn_{0}$>(r,x,p.snd)
}
\end{codejava}
% Note that, to avoid clutter, we have already resolved appropriate
% region arguments to the recursive call, instead of introducing new
% region variables and generating equality constraints on them. 
Constraints generated during the elaboration are shown below
(let $\rhoenv_2 = \{\rhoalloc_0,\rho_{0-4},\rhoalloc_2,\rho_{6-9}\}$ ):
\begin{smathpar}
\begin{array}{l}
\csid{7} \tywf{\rhoenv_2}{\varphi_2}\qquad
\csid{8} \isvalid{\varphi_1 \conj \varphi_2}
    {\rho_7 \outlives \rho_6 \conj \rho_8 \outlives \rho_6} \\
\csid{9} \isvalid{\varphi_1 \conj \varphi_2}{\rho_7 = \rho_9} \qquad
\csid{10} \isvalid{\varphi_1 \conj \varphi_2}{\rho_{8} = \rho_9} \\
\csid{11} \isvalid{\varphi_1 \conj \varphi_2 \conj \rgn_0 \outlives
\rgn_1} 
    {[\rgn_1/\rhoalloc_2][\rgn_1/\rho_6][\rgn_0/\rho_{7-9}]\varphi_2}
\end{array}
\end{smathpar}

@@

\paragraph{Constraint Solution}
Solving the set ($C$) of constraints entails finding an
assignment for each predicate variable $\varphi$) and each region
variable ($\rho$) that occurs free in $C$, such that the solution
satisfies well-formedness constraints on $\varphi$ and $\rhobar$.
% To simplify presentation, we think of $C$ as being parameterized on
% $\varphi$ and $\rhobar$, and write it as $C\lbrack \varphi,\rhobar
% \rbrack$. We now formalize the constraint satisfaction problem, and
% its solution.

\begin{definition}
\emph{(\textbf{Constraint Satisfaction Problem (CSP)})} A constraint
satisfaction problem is a tuple $\CSP$, where $C\lbrack
\varphi,\rhobar \rbrack$ is a set of validity constraints, where
$i$'th validity constraint assumes one of the following forms:
\begin{center}
\(
    \isvalid{\phictxt^{i} \conj \varphi}
            {\phicstr^{i}}\qquad
    \isvalid{\phictxt^{i} \conj \varphi}
            {F_i(\varphi)}
\)
\end{center}
$\bar{\rhoenv}$ are the unification domains for $\rhobar$. We call the
union of all unification domains ($\bigcup\bar{\rhoenv}$) as the
unification domain ($\rhoenv$) of the CSP.  The solution to the
constraint satisfaction problem is a pair $(\substFn,\phisol)$, where
$\substFn$ is a map from $\rhobar$ to $\rhoenv$ such that
$\substFn(\rho_j) \in \rhoenv_j$, for every $j$, and $\phisol$ is a
constraint formula such that:
\begin{itemize}
\item $\phisol$ is well-formed under $\rhoenv_{\varphi}$ (i.e.,
$\tywf{\rhoenv_{\varphi}}{\phisol}$).
\item Every sequent in $C\lbrack (\phisol,\substFn(\rhobar)) \rbrack$
is valid.
% \begin{center}
% \(
%   \isvalid{\phictxt^{i} \conj \bigwedge_j(\rho_j = \substFn(\rho_j)) \conj \phisol}
%           {\phicstr^{i} \conj  F_i(\phisol))}
% \)
% \end{center}
\end{itemize}
\end{definition}

\paragraph{Properties of generated constraints}
The set of generated constraints have certain properties that allow us to
solve the constraints efficiently.
%
Since $\FB$ does not admit null values or uninitialized variables, every region
variable is forced to be unified with some concrete region, by the constraints.

Our approach is based on the observation that $\FB$ does not admit
null values or uninitialized variables, thus forcing every region
variable to be unified with some concrete region.
%
The constraint formula $\Phics$ captures all such unification constraints on $\rhobar$.
Since the set of all concrete region identifiers is the
unification domain ($\rhoenv$) of CSP, this means for every $\rho_i$
with a well-formedness constraint as $\rho_i \in \rhoenv_i$, there
exists a $\rgn \in \rhoenv$ such that and $\isvalid{\Phicx \conj
\Phics}{\rho = \rgn}$. However, $\rgn$ may not belong to $\rhoenv_i$,
in which case the well-formedness constraint on $\rho_i$ is not
satisfied, and constraint solving must fail.  Therefore, there exists
a unique assignment $\substFn$ such that $c\lbrack
\phi_{sol},\substFn(\rhobar)$ is satisfied, regardless of
$\phi_{sol}$.

\item Case $\tynwf{\rhoenv_{\varphi}}{\phi}$: In this case, $\phi$
contains at least one equality or outlives constraint on two region
identifiers, $\rgn_i$ and $\rgn_j$, where (a). $\rgn_i,\rgn_j \in
\rhoenv - \rhoenv_{\varphi}$, or (b). $\rgn_i \in \rhoenv -
\rhoenv_{\varphi}$ and $\rgn_j \in \rhoenv_{\varphi}$, such that the
constraint is not implied by $\Phicx$ (if it is implied, then $\phi$
is not maximally weak). Let us denote such constraint on $\rgn_i$ and
$\rgn_j$ as $\phi_{ij}$. Now, let us consider a solution $\phisol$ to
$\varphi$, which means that $\isvalid{\Phicx \conj \phisol}{\Phics'}$.
Since $\isvalid{\Phicx}{\phi \Leftrightarrow \Phics'}$, we have
$\isvalid {\Phicx \conj \phisol}{\phi}$. Since
$\isvalid{\phi}{\phi_{ij}}$, we have $\isvalid {\Phicx \conj
\phisol}{\phi_{ij}}$, although $\isnotvalid{\Phicx}{\phi_{ij}}$. But
this is impossible. To see why, recall that all the constraints on
identifiers in $\rhoenv_{\varphi}$ occuring in $\Phicx$ are subsumed
by $(\rhoenv - \rhoenv_{\varphi}) \outlives \rhoenv_{\varphi}$.
Therefore, it is impossible to derive $\phi_{ij}$ from $\Phicx$ by
only adding constraints on $\rgn \in \rhoenv_{\varphi}$. Hence, such a
solution $\phisol$ cannot exist.
\end{itemize}

\textbf{TODO: Is the above inequality correct?}

The above discussion hints at an algorithm to compute a solution to
$\varphi$: find a maximally weak $\phisol$ such that $\isvalid{\Phicx}
{\phisol \Leftrightarrow \Phics'}$. If $\tywf{\rhoenv_{\varphi}}
{\phisol}$, then $\phisol$ is the solution. Otherwise, there is no
solution to $\varphi$. We now describe a graph-based algorithm to
compute maximally weak $\phisol$.

\textbf{TODO: FP iteration should use $\Phicx$}

Substituting the above solutions for $\varphi_0$ and $\varphi_1$ in
$c_{14}$ gives us a recursive constraint of the following form:
\begin{center}
\(
  \isvalid{\phictxt \conj \varphi}{\phicstr \conj F(\varphi)}
\)
\end{center}
Where $\phictxt$ and $\phicstr$ are concrete constraint formulas, and
$F$ is a substitution function, not necessarily idempotent. We now
extend constraint solving to recursive CSP $\csp$, where $c\lbrack
\varphi,\rhobar \rbrack$ assumes the form $\isvalid{\Phicx \conj
\varphi}{\Phics \conj \bigwedge_i F_i(\varphi)}$. For the sake of
brevity, we define $G(\varphi) = \Phics \conj \bigwedge_i
F_i(\varphi)$, and use $F$ in $c\lbrack \varphi,\rhobar \rbrack$: 
\begin{center}
\(
  \isvalid{\Phicx \conj \varphi}{G(\varphi)}
\)
\end{center}
To solve the above recursive constraint, we start with the observation
that the set of all constaints ($\phi$) over $\rhoenv \cup
\{\rhobar\}$ is a lattice, where $\phi_1 \le \phi_2 \triangleq
\isvalid{\cdot}{\phi_1 \Rightarrow \phi_2}$. Note that $G$ is a
monotone over the lattice:
\begin{center}
  $\forall \phi_1,\phi_2.\; \isvalid{\cdot}{(\phi_1 \Rightarrow
  \phi_2) \Rightarrow (G(\phi_1) \Rightarrow G(\phi_2))}$
\end{center}
From Knaster-Tarski's theorem, we know that $G$ has a greatest fixed
point ($\phi_f$), with following properties:
\begin{itemize}
\item \textbf{Property 1} $\phi_f = G(\phi_f)$, which means
$\isvalid{\cdot}{\phi_f \Leftrightarrow G(\phi_f)}$
\item \textbf{Property 2} $\forall \phi_f'$ such that $\phi_f' =
G(\phi_f')$, we have $\phi_f' \le \phi_f$, which means $\isvalid
{\cdot} {\phi_f' \Rightarrow \phi_f}$.
\end{itemize}
We therefore compute the greatest fixed point ($\phi_f$) of $G$, and
convert the recursive constraint to the following non-recursive
constraint:
\begin{center}
\(
  \isvalid{\Phicx \conj \varphi}{\phi_f}
\)
\end{center}
The technique described in \S~\ref{sec:csolve} now suffices to solve
the above constraint.

\subsection{Modular Constraint Solving}
\label{sec:fb-constraintsolving}

@@
